---
title: "**S-LDSC analysis to estimate the enrichment of GWAS variants in predicted splicing variants**"
author: "Christie Du"
date: "2020-08-20"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## **Introduction**
I ran S-LDSC on SPIDEX splicing annotations. The SPIDEX database was downloaded from ANNOVAR. I did this for 11 GWAS: Asthma, BMI, CAD, HDL, Height, Intelligence, IBD, LDL, RA, Schizophrenia, and Type II Diabetes. I used BaselineLD v1.1 for S-LDSC analysis.

SPIDEX splicing annotations are located here: `RCC:/project2/xinhe/cdu/ldsc/annot/spidex`
All GWAS in sumstats format are located here: `RCC:/project2/xinhe/cdu/GWAS/ldsc_format`

```{r}
annot_list <- c("splicing")
trait_name_list <- c("Asthma", "BMI","CAD","HDL","Height","IBD","Intelligence","LDL","RA","Schizophrenia","T2D")
```

```{r}
library(ggplot2)
library(foreach)
library(doParallel)

ggplot_enrichment <- function(result_sLDSC, xlim = NULL, title = "Enrichment"){
  
  Enrichment <- result_sLDSC$Enrichment
  Enrichment_CI_L <- result_sLDSC$Enrichment - 1.96*result_sLDSC$Enrichment_std_error
  Enrichment_CI_H <- result_sLDSC$Enrichment + 1.96*result_sLDSC$Enrichment_std_error

  ## truncate at 1
  Enrichment[Enrichment < 1] <- 1
  Enrichment_CI_L[Enrichment_CI_L < 1] <- 1
  Enrichment_CI_H[Enrichment_CI_H < 1] <- 1
  
  p <- ggplot(result_sLDSC, aes(x = Enrichment, y = Disease, colour = Color))+
    geom_point()+
    xlab("Enrichment")+
    ggtitle(title)+
    geom_errorbarh(aes(xmin = Enrichment - 1.96*Enrichment_std_error, 
                       xmax = Enrichment + 1.96*Enrichment_std_error, height = 0.1))+ 
    facet_wrap(Category~.,ncol = 1)+
    theme_bw()  + 
    geom_vline(xintercept = 1,linetype="dotted", colour = "red")+
    theme(axis.ticks = element_blank(),  
          panel.grid.minor = element_blank(), 
          axis.line = element_line(colour = "black"), 
          axis.text = element_text(face="bold",size = 9, colour = "black"),
          axis.title = element_text(face="bold",size = 9),
          strip.text = element_text(face="bold",size = 9), 
          panel.spacing.x = unit(0.6,units = "cm"), 
          axis.title.y = element_blank(), 
          legend.position = "none", 
          plot.title = element_text(hjust = 0.5))
  if(!is.null(xlim)){
    p <- p + coord_cartesian(xlim = xlim)
  }
  print(p)
}

ggplot_log2_enrichment <- function(result_sLDSC, xlim = NULL, title = "Enrichment"){
  
  result_sLDSC$Enrichment_CI_L <- result_sLDSC$Enrichment - 1.96*result_sLDSC$Enrichment_std_error
  result_sLDSC$Enrichment_CI_H <- result_sLDSC$Enrichment + 1.96*result_sLDSC$Enrichment_std_error

  ## truncate at 1
  result_sLDSC$Enrichment[result_sLDSC$Enrichment < 1] <- 1
  result_sLDSC$Enrichment_CI_L[result_sLDSC$Enrichment_CI_L < 1] <- 1
  result_sLDSC$Enrichment_CI_H[result_sLDSC$Enrichment_CI_H < 1] <- 1

  p <- ggplot(result_sLDSC, aes(x = log2(Enrichment), y = Disease, colour = Color))+
    geom_point()+
    xlab("log2(Enrichment)")+
    ggtitle(title)+
    geom_errorbarh(aes(xmin = log2(Enrichment_CI_L), 
                       xmax = log2(Enrichment_CI_H), height = 0.1))+ 
    facet_wrap(Category~.,ncol = 1)+
    theme_bw()  + 
    geom_vline(xintercept = 0,linetype="dotted", colour = "red")+
    theme(axis.ticks = element_blank(),  
          panel.grid.minor = element_blank(), 
          axis.line = element_line(colour = "black"), 
          axis.text = element_text(face="bold",size = 9, colour = "black"),
          axis.title = element_text(face="bold",size = 9),
          strip.text = element_text(face="bold",size = 9), 
          panel.spacing.x = unit(0.6,units = "cm"), 
          axis.title.y = element_blank(), 
          legend.position = "none", 
          plot.title = element_text(hjust = 0.5))
  if(!is.null(xlim)){
    p <- p + coord_cartesian(xlim = xlim)
  }
  print(p)
}

## combine S-LDSC enrichment results across traits
combine_sldsc_traits <- function(trait_name_list, dir_results, baseline, topX, prefix){
  registerDoParallel(cores = 10)
  
  result_sLDSC <- foreach(trait = trait_name_list, .combine = rbind)%dopar%{
    sldsc_results <- read.table(paste0(dir_results,"/", trait, "/", baseline, "/", topX, "/", trait,"_", prefix, "_", baseline, ".results"), header = T, stringsAsFactors = F)
    sldsc.splicing <- sldsc_results[sldsc_results$Category == "L2_1",]
    sldsc.splicing$Category <- "splicing"
    
    sldsc.combined <- rbind(sldsc.splicing)
    sldsc.combined <- cbind(Disease = trait, sldsc.combined)
    sldsc.combined
  }
  return(result_sLDSC)
}
```

```{r}
baseline <- "baselineLDv1.1"
dir_results <- "/project2/xinhe/cdu/ldsc/results/splicing"

#top5
result_sLDSC_top5 <- combine_sldsc_traits(trait_name_list, dir_results, baseline, "top5_filt", "hg19_spidex_0.05_filt")
result_sLDSC_top5$Category <- factor(result_sLDSC_top5$Category, levels = rev(annot_list) )
result_sLDSC_top5$Color <- factor(result_sLDSC_top5$Disease, levels = trait_name_list)

#top2
result_sLDSC_top2 <- combine_sldsc_traits(trait_name_list, dir_results, baseline, "top2_filt", "hg19_spidex_0.02_filt")
result_sLDSC_top2$Category <- factor(result_sLDSC_top2$Category, levels = rev(annot_list) )
result_sLDSC_top2$Color <- factor(result_sLDSC_top2$Disease, levels = trait_name_list)

#top10
result_sLDSC_top10 <- combine_sldsc_traits(trait_name_list, dir_results, baseline, "top10_filt", "hg19_spidex_0.10_filt")
result_sLDSC_top10$Category <- factor(result_sLDSC_top10$Category, levels = rev(annot_list) )
result_sLDSC_top10$Color <- factor(result_sLDSC_top10$Disease, levels = trait_name_list)
```

### Data Tables
```{r}
DT::datatable(format(result_sLDSC_top2[,1:8], digits = 2), options = list(scrollX = TRUE, keys = TRUE),rownames = F, caption = 'Table 1. Top 2%')

DT::datatable(format(result_sLDSC_top5[,1:8], digits = 2), options = list(scrollX = TRUE, keys = TRUE),rownames = F, caption = 'Table 2. Top 5%')

DT::datatable(format(result_sLDSC_top10[,1:8], digits = 2), options = list(scrollX = TRUE, keys = TRUE),rownames = F, caption = 'Table 3. Top 10%')
```

### Enrichment Plots
```{r}
ggplot_enrichment(result_sLDSC_top2, title = "top 2%")
ggplot_enrichment(result_sLDSC_top5, title = "top 5%")
ggplot_enrichment(result_sLDSC_top10, title = "top 10%")
```

### log2(Enrichment) Plots
```{r}
ggplot_log2_enrichment(result_sLDSC_top2, title = "top 2%")
ggplot_log2_enrichment(result_sLDSC_top5, title = "top 5%")
ggplot_log2_enrichment(result_sLDSC_top10, title = "top 10%")
```
